---
title: "Database Schema Reference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Database Schema Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

The bouncer package stores cricket data in a DuckDB database (`bouncer.duckdb`). This vignette documents all tables, their columns, and relationships. The database contains approximately:

- **15,000+ matches** spanning international cricket and major leagues
- **6 million+ deliveries** of ball-by-ball data
- **25,000+ players** in the registry

## Quick Database Access

```{r}
library(bouncer)

# Connect to database (read-only is safest)
conn <- connect_to_bouncer(read_only = TRUE)

# List all tables
DBI::dbListTables(conn)

# Query example
result <- DBI::dbGetQuery(conn, "SELECT COUNT(*) FROM matches")

# Always disconnect when done
disconnect_bouncer(conn)
```

## Table Categories

| Category | Tables | Purpose |
|----------|--------|---------|
| **Core** | matches, deliveries, players, match_innings, innings_powerplays | Raw cricket data |
| **Ratings** | team_elo, player_elo_history | Game-level ELO ratings |
| **Player Skills** | t20/odi/test_player_skill | Ball-by-ball player indices |
| **Team Skills** | t20/odi/test_team_skill | Ball-by-ball team indices |
| **Venue Skills** | t20/odi/test_venue_skill | Venue characteristics |
| **Projections** | t20/odi/test_score_projection | Per-delivery score projections |
| **Predictions** | pre_match_features, pre_match_predictions | Pre-game model data |
| **Simulation** | simulation_results | Monte Carlo outputs |

---

## Core Tables

### matches

Match metadata including teams, venue, outcome, and officials.

| Column | Type | Description |
|--------|------|-------------|
| `match_id` | VARCHAR | **Primary Key** - Unique match identifier (Cricsheet ID) |
| `season` | VARCHAR | Season/year (e.g., "2024", "2023/24") |
| `match_type` | VARCHAR | Format code: T20, IT20, ODI, ODM, Test, MDM |
| `match_type_number` | INTEGER | Match number for Tests (1st, 2nd, etc.) |
| `match_date` | DATE | Start date of match |
| `venue` | VARCHAR | Ground name |
| `city` | VARCHAR | City name |
| `gender` | VARCHAR | "male" or "female" |
| `team_type` | VARCHAR | "international", "club", etc. |
| `team1` | VARCHAR | First team name |
| `team2` | VARCHAR | Second team name |
| `balls_per_over` | INTEGER | Balls per over (usually 6) |
| `overs_per_innings` | INTEGER | Max overs (20 for T20, 50 for ODI, NULL for Test) |
| `toss_winner` | VARCHAR | Team that won the toss |
| `toss_decision` | VARCHAR | "bat" or "field" |
| `outcome_type` | VARCHAR | "normal", "tie", "no result", "draw" |
| `outcome_winner` | VARCHAR | Winning team name |
| `outcome_by_runs` | INTEGER | Margin if won batting first |
| `outcome_by_wickets` | INTEGER | Margin if won chasing |
| `outcome_method` | VARCHAR | "D/L", "VJD" if applicable |
| `umpire1` | VARCHAR | On-field umpire |
| `umpire2` | VARCHAR | On-field umpire |
| `tv_umpire` | VARCHAR | Third umpire |
| `referee` | VARCHAR | Match referee |
| `player_of_match_id` | VARCHAR | Player of match ID |
| `event_name` | VARCHAR | Tournament name (e.g., "Indian Premier League") |
| `event_match_number` | INTEGER | Match number in tournament |
| `event_group` | VARCHAR | Group stage identifier |
| `data_version` | VARCHAR | Cricsheet data format version |
| `data_created` | DATE | When Cricsheet created this file |
| `data_revision` | INTEGER | Cricsheet revision number |

**Example Query:**
```{r}
# Get all IPL 2024 matches
ipl_2024 <- DBI::dbGetQuery(conn, "
  SELECT match_id, match_date, team1, team2, outcome_winner
  FROM matches
  WHERE event_name = 'Indian Premier League' AND season = '2024'
  ORDER BY match_date
")
```

---

### deliveries

Ball-by-ball data - the heart of cricket analytics.

| Column | Type | Description |
|--------|------|-------------|
| `delivery_id` | VARCHAR | **Primary Key** - Format: `{match_id}_{team}_{inn}_{over}_{ball}` |
| `match_id` | VARCHAR | Foreign key to matches |
| `season` | VARCHAR | Season/year |
| `match_type` | VARCHAR | Format code |
| `match_date` | DATE | Match date |
| `venue` | VARCHAR | Ground name |
| `city` | VARCHAR | City |
| `gender` | VARCHAR | "male" or "female" |
| `batting_team` | VARCHAR | Team batting |
| `bowling_team` | VARCHAR | Team bowling |
| `innings` | INTEGER | Innings number (1, 2, 3, 4 for Tests; super overs are 3+) |
| `over` | INTEGER | Over number (0-indexed: 0 = first over) |
| `ball` | INTEGER | Ball within over (1-6, can exceed 6 for extras) |
| `over_ball` | DECIMAL | Decimal format (e.g., 5.3 = 6th over, 3rd ball) |
| `batter_id` | VARCHAR | Batter's player ID |
| `bowler_id` | VARCHAR | Bowler's player ID |
| `non_striker_id` | VARCHAR | Non-striker's player ID |
| `runs_batter` | INTEGER | Runs scored by batter (0-6 typically) |
| `runs_extras` | INTEGER | Extra runs (wides, no balls, byes, leg byes) |
| `runs_total` | INTEGER | Total runs from this delivery |
| `is_boundary` | BOOLEAN | TRUE if boundary (4 or 6) |
| `is_four` | BOOLEAN | TRUE if four |
| `is_six` | BOOLEAN | TRUE if six |
| `wides` | INTEGER | Wide runs (includes +1 for the wide itself) |
| `noballs` | INTEGER | No ball runs |
| `byes` | INTEGER | Bye runs |
| `legbyes` | INTEGER | Leg bye runs |
| `penalty` | INTEGER | Penalty runs |
| `is_wicket` | BOOLEAN | TRUE if wicket fell |
| `wicket_kind` | VARCHAR | Dismissal type (bowled, caught, lbw, etc.) |
| `player_out_id` | VARCHAR | Dismissed player's ID |
| `fielder1_id` | VARCHAR | Primary fielder involved |
| `fielder2_id` | VARCHAR | Secondary fielder (run outs) |
| `fielder1_is_sub` | BOOLEAN | TRUE if fielder was substitute |
| `has_review` | BOOLEAN | TRUE if DRS review taken |
| `review_by` | VARCHAR | Team that reviewed |
| `review_decision` | VARCHAR | "upheld", "overturned", etc. |
| `has_replacement` | BOOLEAN | TRUE if player replacement occurred |
| `replacement_in` | VARCHAR | Incoming player ID |
| `replacement_out` | VARCHAR | Outgoing player ID |
| `replacement_reason` | VARCHAR | "concussion", "impact player", etc. |
| `total_runs` | INTEGER | Running innings total |
| `wickets_fallen` | INTEGER | Running wickets count |

**Delivery ID Format:**
```
{match_id}_{batting_team}_{innings}_{over:03d}_{ball:02d}
Example: "64012_India_1_005_03" = Match 64012, India batting, 1st innings, over 5, ball 3
```

**Example Query:**
```{r}
# Get Kohli's boundaries in IPL 2024
kohli_boundaries <- DBI::dbGetQuery(conn, "
  SELECT d.match_date, d.runs_batter, d.is_four, d.is_six, d.bowler_id
  FROM deliveries d
  JOIN matches m ON d.match_id = m.match_id
  WHERE d.batter_id = 'V Kohli'
    AND d.is_boundary = TRUE
    AND m.event_name = 'Indian Premier League'
    AND m.season = '2024'
")
```

---

### players

Player registry with demographic information.

| Column | Type | Description |
|--------|------|-------------|
| `player_id` | VARCHAR | **Primary Key** - Format varies by source |
| `player_name` | VARCHAR | Full display name |
| `country` | VARCHAR | Primary country |
| `dob` | DATE | Date of birth (where available) |
| `batting_style` | VARCHAR | "right-hand bat", "left-hand bat" |
| `bowling_style` | VARCHAR | "right-arm fast", "slow left-arm orthodox", etc. |

---

### match_innings

Innings-level summaries.

| Column | Type | Description |
|--------|------|-------------|
| `match_id` | VARCHAR | **Primary Key (part)** |
| `innings` | INTEGER | **Primary Key (part)** - Innings number |
| `batting_team` | VARCHAR | Team batting |
| `bowling_team` | VARCHAR | Team bowling |
| `total_runs` | INTEGER | Final innings total |
| `total_wickets` | INTEGER | Wickets lost |
| `total_overs` | DECIMAL | Overs faced |
| `declared` | BOOLEAN | TRUE if innings declared (Tests) |
| `forfeited` | BOOLEAN | TRUE if innings forfeited |
| `target_runs` | INTEGER | Chase target (2nd innings limited overs) |
| `target_overs` | INTEGER | Overs available (may differ due to DLS) |
| `is_super_over` | BOOLEAN | TRUE for super over innings |
| `absent_hurt` | VARCHAR | JSON array of players who couldn't bat |

---

### innings_powerplays

Powerplay periods for limited-overs matches.

| Column | Type | Description |
|--------|------|-------------|
| `powerplay_id` | VARCHAR | **Primary Key** - `{match_id}_{innings}_{seq}` |
| `match_id` | VARCHAR | Match identifier |
| `innings` | INTEGER | Innings number |
| `from_over` | REAL | Powerplay start (e.g., 0.1) |
| `to_over` | REAL | Powerplay end (e.g., 5.6) |
| `powerplay_type` | VARCHAR | "mandatory", "batting", "bowling" |

---

## Rating Tables

### team_elo

Game-level team ELO ratings (updated after each match).

| Column | Type | Description |
|--------|------|-------------|
| `team_id` | VARCHAR | **Primary Key (part)** - Team name |
| `match_id` | VARCHAR | **Primary Key (part)** - Match identifier |
| `match_date` | DATE | Match date |
| `match_type` | VARCHAR | Format code |
| `event_name` | VARCHAR | Tournament name |
| `elo_result` | DOUBLE | Result-based ELO (updates on win/loss/draw) |
| `elo_roster_batting` | DOUBLE | Aggregated batting ELO from player roster |
| `elo_roster_bowling` | DOUBLE | Aggregated bowling ELO from player roster |
| `elo_roster_combined` | DOUBLE | Combined roster ELO |
| `matches_played` | INTEGER | Running match count |

**ELO Interpretation:**
- 1500 = average team
- 1600+ = strong team
- 1400- = weaker team
- Difference of 100 points ≈ 64% expected win rate

---

### player_elo_history

Historical player ELO snapshots (one row per player per match played).

| Column | Type | Description |
|--------|------|-------------|
| `player_id` | VARCHAR | **Primary Key (part)** |
| `match_id` | VARCHAR | **Primary Key (part)** |
| `match_date` | DATE | Match date |
| `match_type` | VARCHAR | Format played |
| `elo_batting` | DOUBLE | Overall batting ELO |
| `elo_bowling` | DOUBLE | Overall bowling ELO |
| `elo_batting_test` | DOUBLE | Test-specific batting ELO |
| `elo_batting_odi` | DOUBLE | ODI-specific batting ELO |
| `elo_batting_t20` | DOUBLE | T20-specific batting ELO |
| `elo_bowling_test` | DOUBLE | Test-specific bowling ELO |
| `elo_bowling_odi` | DOUBLE | ODI-specific bowling ELO |
| `elo_bowling_t20` | DOUBLE | T20-specific bowling ELO |

---

## Player Skill Tables

These tables store per-delivery skill indices using an exponential moving average (EMA) approach. Available for each format: `t20_player_skill`, `odi_player_skill`, `test_player_skill`.

| Column | Type | Description |
|--------|------|-------------|
| `delivery_id` | VARCHAR | **Primary Key** |
| `match_id` | VARCHAR | Match identifier |
| `match_date` | DATE | Match date |
| `batter_id` | VARCHAR | Batter's player ID |
| `bowler_id` | VARCHAR | Bowler's player ID |
| `batter_scoring_index` | DOUBLE | Runs per ball vs expected (+ is better) |
| `batter_survival_rate` | DOUBLE | Probability of surviving each ball (0.95-0.99 typical) |
| `bowler_economy_index` | DOUBLE | Runs conceded vs expected (- is better for bowlers) |
| `bowler_strike_rate` | DOUBLE | Wickets per ball (higher = more wicket-taking) |
| `exp_runs` | DOUBLE | Expected runs (from skill combination) |
| `exp_wicket` | DOUBLE | Expected wicket probability |
| `actual_runs` | INTEGER | Actual runs scored |
| `is_wicket` | BOOLEAN | Whether wicket fell |
| `batter_balls_faced` | INTEGER | Career balls faced (for reliability) |
| `bowler_balls_bowled` | INTEGER | Career balls bowled |

**Skill Index Interpretation:**

| Index | Good Value | Average | Poor Value |
|-------|------------|---------|------------|
| Batter Scoring | +0.10 | 0.00 | -0.10 |
| Batter Survival | 0.99 | 0.97 | 0.95 |
| Bowler Economy | -0.10 | 0.00 | +0.10 |
| Bowler Strike Rate | 0.06 | 0.04 | 0.02 |

---

## Team Skill Tables

Per-delivery team skill indices. Available as `t20_team_skill`, `odi_team_skill`, `test_team_skill`.

| Column | Type | Description |
|--------|------|-------------|
| `delivery_id` | VARCHAR | **Primary Key** |
| `match_id` | VARCHAR | Match identifier |
| `match_date` | DATE | Match date |
| `batting_team_id` | VARCHAR | Batting team |
| `bowling_team_id` | VARCHAR | Bowling team |
| `batting_team_runs_skill` | DOUBLE | Team batting runs residual (+ is better) |
| `batting_team_wicket_skill` | DOUBLE | Team batting wicket residual (+ means loses fewer wickets) |
| `bowling_team_runs_skill` | DOUBLE | Team bowling runs residual (- is better for bowling) |
| `bowling_team_wicket_skill` | DOUBLE | Team bowling wicket residual (+ means takes more wickets) |
| `exp_runs_agnostic` | DOUBLE | Expected runs (agnostic model) |
| `exp_wicket_agnostic` | DOUBLE | Expected wicket prob (agnostic model) |
| `actual_runs` | INTEGER | Actual runs |
| `is_wicket` | BOOLEAN | Wicket fell |
| `batting_team_balls` | INTEGER | Team balls batted (reliability) |
| `bowling_team_balls` | INTEGER | Team balls bowled |

---

## Venue Skill Tables

Per-delivery venue characteristics. Available as `t20_venue_skill`, `odi_venue_skill`, `test_venue_skill`.

| Column | Type | Description |
|--------|------|-------------|
| `delivery_id` | VARCHAR | **Primary Key** |
| `match_id` | VARCHAR | Match identifier |
| `match_date` | DATE | Match date |
| `venue` | VARCHAR | Ground name |
| `venue_run_rate` | DOUBLE | Average runs per ball at venue |
| `venue_wicket_rate` | DOUBLE | Average wicket probability at venue |
| `venue_boundary_rate` | DOUBLE | Boundary percentage at venue |
| `venue_dot_rate` | DOUBLE | Dot ball percentage at venue |
| `venue_balls` | INTEGER | Total balls at venue (reliability) |
| `actual_runs` | INTEGER | Actual runs this delivery |
| `is_wicket` | BOOLEAN | Wicket fell |
| `is_boundary` | BOOLEAN | Boundary hit |
| `is_dot` | BOOLEAN | Dot ball |

---

## Score Projection Tables

Per-delivery innings projections. Available as `t20_score_projection`, `odi_score_projection`, `test_score_projection`.

| Column | Type | Description |
|--------|------|-------------|
| `delivery_id` | VARCHAR | **Primary Key** |
| `match_id` | VARCHAR | Match identifier |
| `match_date` | DATE | Match date |
| `innings` | INTEGER | Innings number |
| `batting_team_id` | VARCHAR | Batting team |
| `current_score` | INTEGER | Score at this delivery |
| `balls_remaining` | INTEGER | Balls left in innings |
| `wickets_remaining` | INTEGER | Wickets in hand |
| `resource_remaining` | DOUBLE | DLS-style resource % (0-1) |
| `resource_used` | DOUBLE | Resources consumed (0-1) |
| `eis_agnostic` | DOUBLE | Expected Initial Score (baseline model) |
| `eis_full` | DOUBLE | Expected Initial Score (full model with skills) |
| `projected_agnostic` | DOUBLE | Projected final score (baseline) |
| `projected_full` | DOUBLE | Projected final score (with skills) |
| `final_innings_total` | INTEGER | Actual final score (for validation) |
| `projection_change_agnostic` | DOUBLE | Change from previous delivery |
| `projection_change_full` | DOUBLE | Change from previous delivery |

---

## Prediction Tables

### pre_match_features

Features used for pre-match prediction models.

| Column | Type | Description |
|--------|------|-------------|
| `match_id` | VARCHAR | **Primary Key** |
| `match_date` | DATE | Match date |
| `match_type` | VARCHAR | Format code |
| `event_name` | VARCHAR | Tournament |
| `team1` | VARCHAR | First team |
| `team2` | VARCHAR | Second team |
| `team1_elo_result` | DOUBLE | Team 1's result-based ELO |
| `team1_elo_roster` | DOUBLE | Team 1's roster-based ELO |
| `team1_form_last5` | DOUBLE | Team 1's win rate in last 5 matches |
| `team1_h2h_wins` | INTEGER | Team 1's H2H wins vs Team 2 |
| `team1_h2h_total` | INTEGER | Total H2H matches |
| `team1_bat_scoring_avg` | DOUBLE | Avg batting scoring index |
| `team1_bat_scoring_top5` | DOUBLE | Top 5 batters' avg scoring index |
| `team1_bat_survival_avg` | DOUBLE | Avg survival rate |
| `team1_bowl_economy_avg` | DOUBLE | Avg bowling economy index |
| `team1_bowl_economy_top5` | DOUBLE | Top 5 bowlers' avg economy |
| `team1_bowl_strike_avg` | DOUBLE | Avg bowling strike rate |
| `team2_*` | ... | Same features for Team 2 |
| `venue` | VARCHAR | Match venue |
| `venue_avg_score` | DOUBLE | Average first innings score at venue |
| `venue_chase_success_rate` | DOUBLE | Historical chase success rate |
| `venue_matches` | INTEGER | Matches at venue (reliability) |
| `is_knockout` | BOOLEAN | Knockout/playoff match |
| `is_neutral_venue` | BOOLEAN | Neutral venue |
| `expected_margin` | DOUBLE | Predicted margin from ELO |
| `actual_margin` | DOUBLE | Actual margin (backfilled) |

---

### pre_match_predictions

Model predictions for each match.

| Column | Type | Description |
|--------|------|-------------|
| `prediction_id` | VARCHAR | **Primary Key** |
| `match_id` | VARCHAR | Match identifier |
| `model_version` | VARCHAR | Model version string |
| `model_type` | VARCHAR | Model architecture |
| `prediction_date` | TIMESTAMP | When prediction was made |
| `team1_win_prob` | DOUBLE | Team 1 win probability (0-1) |
| `team2_win_prob` | DOUBLE | Team 2 win probability (0-1) |
| `predicted_winner` | VARCHAR | Predicted winning team |
| `confidence` | DOUBLE | Model confidence |
| `actual_winner` | VARCHAR | Actual winner (backfilled) |
| `prediction_correct` | BOOLEAN | Was prediction correct |

---

### simulation_results

Monte Carlo simulation outputs.

| Column | Type | Description |
|--------|------|-------------|
| `simulation_id` | VARCHAR | **Primary Key** |
| `simulation_type` | VARCHAR | "season", "playoffs", "match" |
| `event_name` | VARCHAR | Tournament simulated |
| `season` | VARCHAR | Season |
| `simulation_date` | TIMESTAMP | When simulation ran |
| `n_simulations` | INTEGER | Number of iterations |
| `parameters` | VARCHAR | JSON simulation parameters |
| `team_results` | VARCHAR | JSON team-level results |
| `match_results` | VARCHAR | JSON match-level results |
| `created_at` | TIMESTAMP | Creation timestamp |

---

## Parameter Tables

### projection_params

Optimized parameters for score projection by segment.

| Column | Type | Description |
|--------|------|-------------|
| `segment_id` | VARCHAR | **Primary Key** - Format/gender/team_type combo |
| `format` | VARCHAR | t20, odi, test |
| `gender` | VARCHAR | male, female |
| `team_type` | VARCHAR | international, club |
| `param_a` | DOUBLE | Resource curve parameter |
| `param_b` | DOUBLE | Resource curve parameter |
| `param_z` | DOUBLE | Wicket decay parameter |
| `param_y` | DOUBLE | Additional parameter |
| `eis_agnostic` | DOUBLE | Expected initial score (format average) |
| `train_rmse` | DOUBLE | Training RMSE |
| `validation_rmse` | DOUBLE | Validation RMSE |
| `n_innings` | INTEGER | Innings in training set |

---

## Helper Functions

The bouncer package provides functions to load these tables into R without writing SQL:

```{r}
# Load core tables
matches <- load_matches(format = "t20", season = "2024")
deliveries <- load_deliveries(match_ids = "1234567")
players <- load_players()

# Load skill tables
player_skills <- load_player_skill(format = "t20")
team_skills <- load_team_skill(format = "t20")
venue_skills <- load_venue_skill(format = "t20")

# Load ELO ratings
team_elo <- load_team_elo(team = "India", format = "t20")
```

## Live Database Statistics

Run this code to see current database statistics:

```{r}
library(bouncer)

conn <- connect_to_bouncer(read_only = TRUE)

# Table row counts
tables <- c("matches", "deliveries", "players", "match_innings",
            "team_elo", "t20_player_skill", "t20_team_skill", "t20_venue_skill")

for (tbl in tables) {
  count <- DBI::dbGetQuery(conn, sprintf("SELECT COUNT(*) as n FROM %s", tbl))
  cat(sprintf("%s: %s rows\n", tbl, format(count$n, big.mark = ",")))
}

disconnect_bouncer(conn)
```

## See Also

- `vignette("getting-started")` - Package setup and basic usage
- `vignette("player-analysis")` - Working with player skill indices
- `vignette("match-analysis")` - Match-level analysis
- `vignette("predictions")` - Using prediction models
