---
title: "Match Predictions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Match Predictions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction
This vignette covers bouncer's prediction capabilities:
- **Pre-match predictions**: Who's favored before the game starts?
- **In-match win probability**: Live updates as the game progresses
- **Score projections**: What's the expected final score?

## Pre-Match Predictions

Predict the outcome before a match begins:

```{r}
library(bouncer)

# Basic match prediction
prediction <- predict_match("India", "Australia", format = "t20")
print(prediction)
```

The prediction uses:
- Team ELO ratings
- Recent form
- Head-to-head history
- Team skill indices

### Adding Venue Context

```{r}
# Prediction with venue
prediction <- predict_match(
  team1 = "India",
  team2 = "Australia",
  format = "t20",
  venue = "Melbourne Cricket Ground"
)
print(prediction)
```

## In-Match Win Probability

Calculate live win probability during a match:

```{r}
# First innings: batting team at 85/2 after 10 overs
wp <- predict_win_probability(
  current_score = 85,
  wickets = 2,
  overs = 10.0,
  innings = 1,
  format = "t20"
)
print(wp)
```

### Second Innings (Chasing)

```{r}
# Second innings: chasing 180, currently 100/3 after 12.4 overs
wp <- predict_win_probability(
  current_score = 100,
  wickets = 3,
  overs = 12.4,
  innings = 2,
  target = 180,
  format = "t20"
)
print(wp)

# The required run rate is factored into the probability
```

## Score Projection

Project the final innings score from any game state:

```{r}
# Scoreboard: "Mumbai Indians 80/3 (10.0 overs)" in T20
projected <- calculate_projected_score(
  current_score = 80,
  wickets = 3,        # wickets fallen (matches scoreboard)
  overs = 10.0,       # overs bowled (matches scoreboard)
  format = "t20"
)
print(projected)
```

### Understanding Resource Percentage

The projection uses a resource-based formula similar to Duckworth-Lewis:

```{r}
# Calculate resource percentage remaining
resource <- calculate_projection_resource(
  wickets_remaining = 7,  # 10 - 3 wickets fallen
  balls_remaining = 60,   # 10 overs left
  format = "t20"
)
print(resource)  # e.g., 0.65 = 65% of resources remaining
```

Resources depend on:
- Wickets in hand (more wickets = more resources)
- Balls remaining (more balls = more resources)
- The interaction between them

## Understanding ELO and Win Probability

The `calculate_expected_outcome()` function shows how ELO ratings convert to win probability:

```{r}
# Educational: How does ELO work?
# Team A has 1650 ELO, Team B has 1550 ELO
prob_a_wins <- calculate_expected_outcome(1650, 1550)
print(prob_a_wins)  # ~0.64 (64% chance for Team A)

# Equal teams
prob_equal <- calculate_expected_outcome(1500, 1500)
print(prob_equal)  # 0.50 (50-50)

# Large ELO gap
prob_favorite <- calculate_expected_outcome(1700, 1400)
print(prob_favorite)  # ~0.85 (85% chance for higher-rated team)
```

The formula is: `P(A wins) = 1 / (1 + 10^((ELO_B - ELO_A) / 400))`

## Matchup Predictions

Predict specific batter vs bowler outcomes:

```{r}
# What happens when Kohli faces Bumrah?
matchup <- predict_matchup_outcome(
  batter = "V Kohli",
  bowler = "J Bumrah",
  format = "t20"
)
print(matchup)
```

## Team Comparison

Get a comprehensive comparison of two teams:

```{r}
comparison <- compare_teams("India", "Australia", format = "t20")
print(comparison)

# Includes:
# - Current ELO ratings
# - Win probability
# - Recent form
# - Head-to-head record
# - Skill comparisons
```

## Prediction Accuracy

The prediction models are calibrated on historical data. Typical accuracy:

| Prediction Type | Accuracy |
|-----------------|----------|
| Pre-match winner | ~65-70% |
| Score projection | RMSE ~15-20 runs |
| Win probability | Well-calibrated (50% events happen ~50% of time) |

## See Also

- `vignette("getting-started")` - Package overview and setup
- `vignette("simulation")` - Monte Carlo match simulation
- `vignette("match-analysis")` - Analyzing completed matches
