---
title: "Match Simulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Match Simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

This vignette covers bouncer's Monte Carlo simulation capabilities. By simulating matches thousands of times, you can:
- Estimate win probabilities with uncertainty ranges
- Understand the distribution of possible outcomes
- Simulate entire seasons to calculate playoff odds

## Why Simulate?

A single prediction gives one number (e.g., "65% win probability"). Simulation gives you the full distribution: not just who wins, but by how much, and how often various scenarios occur.

## Match Simulation

Simulate a single match many times:

```{r}
library(bouncer)

# Simulate India vs Australia 10,000 times
results <- simulate_match_ballbyball(
  team1 = "India",
  team2 = "Australia",
  format = "t20",
  n_sims = 10000
)

# View summary
print(results)
```

The results include:
- Win percentage for each team
- Score distribution (mean, median, percentiles)
- Margin distribution

## Quick Match Simulation

For faster results with less detail:

```{r}
# Quick simulation (doesn't simulate ball-by-ball)
quick_results <- quick_match_simulation(
  team1 = "India",
  team2 = "Australia",
  format = "t20"
)
print(quick_results)
```

## Simulation Configuration

Customize simulation parameters:

```{r}
# Create custom simulation config
config <- create_simulation_config(
  n_sims = 5000,
  seed = 42,  # For reproducibility
  format = "t20"
)

# Use in simulation
results <- simulate_match_ballbyball(
  team1 = "Mumbai Indians",
  team2 = "Chennai Super Kings",
  format = "t20",
  config = config
)
```

## Innings Simulation

Simulate a single innings:

```{r}
# Simulate an innings
innings_result <- simulate_innings(
  batting_team = "India",
  bowling_team = "Australia",
  format = "t20",
  target = NULL  # First innings, no target
)
print(innings_result)

# Simulate a chase
chase_result <- simulate_innings(
  batting_team = "Australia",
  bowling_team = "India",
  format = "t20",
  target = 185
)
print(chase_result)
```

## Season Simulation

Simulate an entire tournament:

```{r}
# Simulate IPL season
season_results <- simulate_season(
  event = "Indian Premier League",
  season = "2024",
  n_sims = 1000
)

# View playoff probabilities
print(season_results$playoff_odds)
```

### Running Multiple Seasons

For more accurate playoff odds:

```{r}
# Run 10,000 season simulations
season_n <- simulate_season_n(
  event = "Indian Premier League",
  season = "2024",
  n_sims = 10000
)

# Each team's probability of:
# - Making playoffs
# - Finishing top 2
# - Winning the tournament
print(season_n$summary)
```

## IPL Playoff Simulation

Simulate the IPL playoff bracket specifically:

```{r}
# Simulate playoffs with the top 4 teams
playoff_results <- simulate_ipl_playoffs(
  qualifier1_teams = c("Gujarat Titans", "Chennai Super Kings"),
  eliminator_teams = c("Lucknow Super Giants", "Mumbai Indians"),
  n_sims = 10000,
  format = "t20"
)

# Championship probabilities
print(playoff_results)
```

## Interpreting Results

### Win Probability Confidence

From 10,000 simulations:
- If Team A wins 6,500 simulations, we're confident they win ~65% of the time
- The uncertainty is roughly ±1% for 10,000 sims

### Score Distribution

```{r}
# Example output interpretation
# Mean score: 165.3
# Median score: 164
# 10th percentile: 142 (bad day)
# 90th percentile: 188 (good day)
```

### Margin Distribution

Simulations also show likely winning margins:
- Median winning margin
- Probability of close games (< 10 runs or < 2 wickets)
- Probability of blowouts

## Setting Seeds for Reproducibility

```{r}
# Same seed = same results
set.seed(42)
result1 <- simulate_match_ballbyball("India", "Australia", "t20", n_sims = 100)

set.seed(42)
result2 <- simulate_match_ballbyball("India", "Australia", "t20", n_sims = 100)

# result1 and result2 are identical
```

## Computational Considerations

| Simulations | Time (typical) | Use Case |
|-------------|----------------|----------|
| 100 | < 1 second | Quick check |
| 1,000 | ~5 seconds | Reasonable estimate |
| 10,000 | ~30 seconds | Publication quality |
| 100,000 | ~5 minutes | High precision research |

## See Also

- `vignette("getting-started")` - Package overview and setup
- `vignette("predictions")` - Single-point predictions
- `vignette("match-analysis")` - Analyzing real match data
